---
title: Advanced Topics in Machine Learning
subtitle: Sheet 7
author: Submitted by - Ranji Raj
date: "`r Sys.Date()`"
output:
  pdf_document: default
  word_document: default
  html_document:
    df_print: paged
header-includes:
- \usepackage{titling}
- \posttitle{\end{center}} \usepackage{fancyhdr} \pagestyle{fancy} 
---

# **Assignment 7.2 - Branch & Bound (BB)**

__Consider the following given data points in__ $R^2$ __: (1; 3) and (2; 4) labeled with (+) (3; 1) and (4; 2) labeled with (-); (1; 1), (1; 2), (4; 3), (4; 4) unlabeled. Show how the branch and bound method would determine the best hyperplane. For this purpose, visualize the search tree and the corresponding hyperplane for each node. You do not need to exactly compute the values for the minimization. A roughly estimated hyperplane in a diagram is sufficient.__

```{r, echo=FALSE, out.width = '70%', fig.align="center", warning=F, message=F}
library(tidyverse)

df <- data.frame(x = c(1,2,3,4,1,1,4,4)
                 ,y = c(3,4,1,2,1,2,3,4))
                 

df_arrow <- data.frame(x = c(0, 0),
                       y = c(0, 0),
                       xend = c(0, 8),
                       yend = c(8, 0)) 

top_line <- data.frame(x = c(0,1,2,3,4,5,6,7),
                       y = c(1,2,3,4,5,6,7,8))

bottom_line <- data.frame(x = c(1,2,3,4,5,6,7,8),
                          y = c(0,1,2,3,4,5,6,7))

df %>% 
  dplyr::mutate(z = ifelse(x > y + 1, "a",
                           ifelse(x < y - 1, "b", "c"))) %>%
  ggplot(aes(x = x, y = y)) +
  geom_point(aes(shape = z, color = z), size = 5) +
  scale_x_continuous(breaks = 1:7, expand = expansion(add = c(0, 1)))+
  scale_y_continuous(breaks = 1:7, expand = expansion(add = c(0, 1)))+
  coord_fixed(xlim = c(0, 7), ylim = c(0, 7), clip = "off")+
  geom_segment(data = df_arrow, aes(x = x, xend = xend, y = y, yend = yend), size = 0.75, colour = "black",
               arrow = arrow(angle = 20, length = unit(3, "mm"), ends = "last", type = "closed"), linejoin = "mitre") +
  annotate("text", x = c(7.8, 0.3), y = c(0.3, 7.8), label = c("italic(x)", "italic(y)"), parse = TRUE, size = 6)+
  labs(x = NULL,
       y = NULL)+
  theme_bw()+
  theme(panel.grid.major = element_line(colour = "gray80"),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.ticks.length = unit(1, "mm"),
        text = element_text(size = 18),
        legend.position = "none") +
  scale_colour_manual(values = c('red', 'blue', 'black')) +
  scale_shape_manual(breaks = c("a", "b", "c"),
                     values = c("a" = 95, "b" = 3, "c" = 19))
```
Root node: Nothing in $X_u$ labeled; Search tree

```{r , echo=FALSE, out.width = '70%', fig.align="center", warning=F, message=F}
library(data.tree)

acme <- Node$new("root:Unlabeled")
  #accounting <- acme$AddChild("Accounting")

print(acme)
```

Now, greedily select (1; 2) and classify as (+)

```{r, echo=FALSE, out.width = '70%', fig.align="center", warning=F, message=F}
library(tidyverse)

df <- data.frame(x = c(1,2,3,4,1,1,4,4)
                 ,y = c(3,4,1,2,1,2,3,4))
                 

df_arrow <- data.frame(x = c(0, 0),
                       y = c(0, 0),
                       xend = c(0, 8),
                       yend = c(8, 0)) 

top_line <- data.frame(x = c(0,1,2,3,4,5,6,7),
                       y = c(1,2,3,4,5,6,7,8))

bottom_line <- data.frame(x = c(2,3,4,5,6,7,8,9),
                          y = c(0,1,2,3,4,5,6,7))

df %>% 
  dplyr::mutate(z = ifelse(x > y + 1, "a",
                           ifelse(x <= y - 1, "b", "c"))) %>%
  ggplot(aes(x = x, y = y)) +
  geom_point(aes(shape = z, color = z), size = 5) +
  geom_line(data = top_line, aes(x = x, y = y), col="blue") +
  geom_abline(aes(slope = 1, intercept = 0)) +
  geom_line(data = bottom_line, aes(x = x, y = y), col="red") +
  scale_x_continuous(breaks = 1:7, expand = expansion(add = c(0, 1)))+
  scale_y_continuous(breaks = 1:7, expand = expansion(add = c(0, 1)))+
  coord_fixed(xlim = c(0, 7), ylim = c(0, 7), clip = "off")+
  geom_segment(data = df_arrow, aes(x = x, xend = xend, y = y, yend = yend), size = 0.75, colour = "black",
               arrow = arrow(angle = 20, length = unit(3, "mm"), ends = "last", type = "closed"), linejoin = "mitre") +
  annotate("text", x = c(7.8, 0.3), y = c(0.3, 7.8), label = c("italic(x)", "italic(y)"), parse = TRUE, size = 6)+
  labs(x = NULL,
       y = NULL)+
  theme_bw()+
  theme(panel.grid.major = element_line(colour = "gray80"),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.ticks.length = unit(1, "mm"),
        text = element_text(size = 18),
        legend.position = "none") +
  scale_colour_manual(values = c('red', 'blue', 'black')) +
  scale_shape_manual(breaks = c("a", "b", "c"),
                     values = c("a" = 95, "b" = 3, "c" = 19))
```

```{r , echo=FALSE, out.width = '70%', fig.align="center", warning=F, message=F}
library(data.tree)

acme <- Node$new("root:Unlabeled")
  accounting <- acme$AddChild("(1;2)")
    software <- accounting$AddChild("(1;2)+")
print(acme)
```

Now, greedily select (1; 1) and classify as (+)

```{r, echo=FALSE, out.width = '70%', fig.align="center", warning=F, message=F}
library(data.tree)

acme <- Node$new("root:Unlabeled")
  accounting <- acme$AddChild("(1;2)")
    standards <- accounting$AddChild("(1;2)+")
    newLabs <- standards$AddChild("(1;1)+")
  

print(acme)
```
Likewise we select (4;4) as (+) and (4;3) as (-).

The probably approximate hyperplane would be-
```{r, echo=FALSE, out.width = '70%', fig.align="center", warning=F, message=F}
library(tidyverse)

df <- data.frame(x = c(1,2,3,4,1,1,4,4)
                 ,y = c(3,4,1,2,1,2,3,4))
                 

df_arrow <- data.frame(x = c(0, 0),
                       y = c(0, 0),
                       xend = c(0, 8),
                       yend = c(8, 0)) 

top_line <- data.frame(x = c(0,1,2,3,4,5,6,7),
                       y = c(1,2,3,4,5,6,7,8))

bottom_line <- data.frame(x = c(1,2,3,4,5,6,7,8),
                          y = c(0,1,2,3,4,5,6,7))

df %>% 
  dplyr::mutate(z = ifelse(x-1 >= y, "a",
                           ifelse(x < y , "b", "b"))) %>%
  ggplot(aes(x = x, y = y)) +
  geom_point(aes(shape = z, color = z), size = 5) +
  geom_line(data = top_line, aes(x = x, y = y), col="blue") +
  geom_abline(aes(slope = 1, intercept = 0)) +
  geom_line(data = bottom_line, aes(x = x, y = y), col="red") +
  scale_x_continuous(breaks = 1:7, expand = expansion(add = c(0, 1)))+
  scale_y_continuous(breaks = 1:7, expand = expansion(add = c(0, 1)))+
  coord_fixed(xlim = c(0, 7), ylim = c(0, 7), clip = "off")+
  geom_segment(data = df_arrow, aes(x = x, xend = xend, y = y, yend = yend), size = 0.75, colour = "black",
               arrow = arrow(angle = 20, length = unit(3, "mm"), ends = "last", type = "closed"), linejoin = "mitre") +
  annotate("text", x = c(7.8, 0.3), y = c(0.3, 7.8), label = c("italic(x)", "italic(y)"), parse = TRUE, size = 6)+
  labs(x = NULL,
       y = NULL)+
  theme_bw()+
  theme(panel.grid.major = element_line(colour = "gray80"),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.ticks.length = unit(1, "mm"),
        text = element_text(size = 18),
        legend.position = "none") +
  scale_colour_manual(values = c('red', 'blue')) +
  scale_shape_manual(breaks = c("a", "b"),
                     values = c("a" = 95, "b" = 3))
```
The above hyperplane gives a decent separation between the positive and negative region and probably with a minimum objective function.

```{r, echo=FALSE, out.width = '70%', fig.align="center", warning=F, message=F}
library(data.tree)

acme <- Node$new("root:Unlabeled")
  accounting <- acme$AddChild("(1;2)")
    standards <- accounting$AddChild("(1;2)+")
    newLabs <- standards$AddChild("(1;1)+")
    newLabs1 <- newLabs$AddChild("(4;4)+")
    newLabs2 <- newLabs1$AddChild("(4;3)-")

print(acme)
```


### Complete Search space tree


```{r, echo=FALSE, out.width = '70%', fig.align="center", warning=F, message=F}
library(data.tree)

acme <- Node$new("root:Unlabeled")
  accounting <- acme$AddChild("(1;2)")
    software <- accounting$AddChild("(1;2)+")
    standards <- accounting$AddChild("(1;2)-")
  accounting <- acme$AddChild("(1;1)")
    software <- accounting$AddChild("(1;1)+")
    standards <- accounting$AddChild("(1;1)-")
    
    research1 <- acme$AddChild("(4;4)")
    newProductLine <- research1$AddChild("(4;4)+")
    newLabs <- research1$AddChild("(4;4)-")
  research2 <- acme$AddChild("(4;3)")
    newProductLine <- research2$AddChild("(4;3)+")
    newLabs <- research2$AddChild("(4;3)-")

print(acme)
```



```{r, message=F, echo=F, warning=F}
library(fontawesome)
```

\centering

[`r fa("github", fill = "black")`](https://github.com/ranjiGT/ATiML-amendments)