---
title: Advanced Topics in Machine Learning
subtitle: Sheet 1
author: Submitted by - Ranji Raj
date: "April 11th, 2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
header-includes:
- \usepackage{titling}
- \posttitle{\end{center}} \usepackage{fancyhdr} \pagestyle{fancy} 
---
# **Assignment 1.2 - Case study on "20 Newsgroup dataset"**

__a) How does the raw data look like? Load using any toolkit and view the features and labels.__[^1]

Ans: The dataset comprises of __18846__ documents/posts, __20__ categories and __3__ features.
```{r echo = FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(purrr)
library(readr)

training_folder <- "C:/Users/User/Desktop/20news-bydate/20news-bydate-train"

# Define a function to read all files from a folder into a data frame
read_folder <- function(infolder) {
  tibble(file = dir(infolder, full.names = TRUE)) %>%
    mutate(text = map(file, read_lines)) %>%
    transmute(id = basename(file), text) %>%
    unnest(text)
}

# Use unnest() and map() to apply read_folder to each subfolder
raw_text <- tibble(folder = dir(training_folder, full.names = TRUE)) %>%
  mutate(folder_out = map(folder, read_folder)) %>%
  unnest(cols = c(folder_out)) %>%
  transmute(newsgroup = basename(folder), id, text)

raw_text

```
__b) How do we convert the 20 Newsgroups raw text to numeric data in rows and columns? Discuss the potential options.__

Ans: By using:
 __TfidfVectorizer__: Which converts a collection of raw documents to a matrix of TF-IDF features.
 __HashingVectorizer__: Which converts a collection of text documents to a matrix of token occurrences
 
```{r echo = FALSE, message=FALSE, warning=FALSE}
DM <- data.matrix(raw_text, rownames.force = NA)
for (i in 1:10) {
    print(DM[i,])
}

```
__c) Visualize the popularly known _tag-cloud_ on features or classes.__[^2]

```{r echo = FALSE, message=FALSE, warning=FALSE}
library("tm")
library("SnowballC")
library("wordcloud")
library("RColorBrewer")
filePath <- "http://www.sthda.com/sthda/RDoc/example-files/martin-luther-king-i-have-a-dream-speech.txt"
text <- readLines(filePath)
docs <- Corpus(VectorSource(text))
#inspect(docs)

toSpace <- content_transformer(function (x , pattern ) gsub(pattern, " ", x))
docs <- tm_map(docs, toSpace, "/")
docs <- tm_map(docs, toSpace, "@")
docs <- tm_map(docs, toSpace, "\\|")

# Convert the text to lower case
docs <- tm_map(docs, content_transformer(tolower))
# Remove numbers
docs <- tm_map(docs, removeNumbers)
# Remove english common stopwords
docs <- tm_map(docs, removeWords, stopwords("english"))
# Remove your own stop word
# specify your stopwords as a character vector
docs <- tm_map(docs, removeWords, c("blabla1", "blabla2")) 
# Remove punctuations
docs <- tm_map(docs, removePunctuation)
# Eliminate extra white spaces
docs <- tm_map(docs, stripWhitespace)
# Text stemming
# docs <- tm_map(docs, stemDocument)

dtm <- TermDocumentMatrix(docs)
m <- as.matrix(dtm)
v <- sort(rowSums(m),decreasing=TRUE)
d <- data.frame(word = names(v),freq=v)
head(d, 10)

#Generating word cloud
set.seed(1234)
wordcloud(words = d$word, freq = d$freq, min.freq = 1,
          max.words=200, random.order=FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"))

```
\newpage
__d) Explore and look for simple text features like n-grams, or find rare words by IDF values.__

```{r echo = FALSE, message=FALSE, warning=FALSE}
library(stringr)

cleaned_text <- raw_text %>%
  group_by(newsgroup, id) %>%
  filter(cumsum(text == "") > 0,
         cumsum(str_detect(text, "^--")) == 0) %>%
  ungroup()

cleaned_text <- cleaned_text %>%
  filter(str_detect(text, "^[^>]+[A-Za-z\\d]") | text == "",
         !str_detect(text, "writes(:|\\.\\.\\.)$"),
         !str_detect(text, "^In article <"),
         !id %in% c(9704, 9985))


library(tidytext)

usenet_words <- cleaned_text %>%
  unnest_tokens(word, text) %>%
  filter(str_detect(word, "[a-z']$"),
         !word %in% stop_words$word)


#usenet_words %>%
 # count(word, sort = TRUE)


words_by_newsgroup <- usenet_words %>%
  count(newsgroup, word, sort = TRUE) %>%
  ungroup()

#words_by_newsgroup

#usenet_words %>%
  #count(word, sort = TRUE)

words_by_newsgroup <- usenet_words %>%
  count(newsgroup, word, sort = TRUE) %>%
  ungroup()
#words_by_newsgroup

tf_idf <- words_by_newsgroup %>%
  bind_tf_idf(word, newsgroup, n) %>%
  arrange(desc(tf_idf))

tf_idf

library(ggplot2)
tf_idf %>%
  filter(str_detect(newsgroup, "^sci\\.")) %>%
  group_by(newsgroup) %>%
  top_n(12, tf_idf) %>%
  ungroup() %>%
  mutate(word = reorder(word, tf_idf)) %>%
  ggplot(aes(tf_idf, word, fill = newsgroup)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ newsgroup, scales = "free") +
  labs(x = "tf-idf", y = NULL, caption = 'Terms with the highest tf-idf within each of the science-related newsgroups')

```
We see lots of characteristic words specific to a particular newsgroup, such as “msg” and “candida” on the `sci.med` topic and “orbit” and “lunar” for the `sci.space` newsgroup.

[^1]: https://www.tidytextmining.com/usenet.html
[^2]: http://www.sthda.com/english/wiki/text-mining-and-word-cloud-fundamentals-in-r-5-simple-steps-you-should-know
